---
title: "101a week 5"
author: "Joshua Susanto"
date: "4/26/2022"
output: html_document
---

```{r}
library('tidyverse')
loans <- read_csv('collegeloans.csv')
head(loans)
public <- loans %>% filter(type == 'public')
private <- loans %>% filter(type == 'private NFP')
unique(loans$type)
plot(public$loan_avg_amount)
plot(private$loan_avg_amount)
```

```{r}
temp <- if_else(loans$type == 'public', 1 ,0)
loans$is_public <- temp
head(loans)
boxplot(loans$loan_avg_amount  ~ loans$is_public)
```

mean function
E(Y|X) --> what is the mean value of y when x is something

```{r}
plot(loans$loan_avg_amount  ~ loans$is_public)
m1 <- lm(loans$loan_avg_amount  ~ loans$is_public)
abline(m1)
```
^ plots the means together

```{r}
summary(m1)
plot(m1)
```
we would get the mean for the second value by just addint the slope to the intercept one time

```{r}
all <- read_csv("collegeloansall.csv")
table(all$type)
m2 <- lm(all$loan_avg_amount ~ all$type)
summary(m2) #everything is being compared to the first intercpt variable
boxplot(all$loan_avg_amount ~ all$type)
```
can also use factors to make sure that it treated properly
lm(y ~ factor(x))

there will always be small variability in the qq plot so focus on the center and make sure there isnt too large of a deviance from the line and make sure there are no patterns

outliers --> points that dont fit the trend
leverage --> a point that will greatly change the model value if its y value is changed (usually high x variance)
influential --> will

if the fitted line goes through an outlier --> high leverage


outlier x value --> high leverage


good v bad leverage
- good --> follows trend
- bad --> does not, outlier
  - no reason for the line being the way it is other than the fact that the leverage point is affecting it










