---
title: "stats 101a week 1"
author: "Joshua Susanto"
date: "3/28/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
rain <- read.csv('rain-1.csv')
head(rain)
library('tidyverse')
library('ggplot2')
library('dplyr')
library('knitr')
```
# list of what is in a dataset
outliers
number of obs 54057
number of vars
types of variables
are there missing values
how was this data collected
what do the variables mean
where are the locations -> how many are there?
units of measurement
timespan
size of dataset (gb mb)

```{r}
str(rain) #tells you the type of variable for each colimn
dim(rain) #gives you the shape of the dataset
# is.na(rain) #lets you know which obervations are na
sum(is.na(rain))
```

data moves

#from dplyr
filter(rain) #filters out rows you specify 
arrange(rain)
select()
mutate()
group_by()
summarize()


```{r}
?filter
rain1 <- filter(rain, PRCP >= 1)
dim(rain1)
```

```{r}
rain2 <- filter(rain, NAME == 'CULVER CITY, CA US')
dim(rain2)
```

```{r}
?arrange
rain3 <- arrange(rain, desc(PRCP)) #desc = descending
a <- max(rain3$PRCP,na.rm = TRUE)
```

```{r}
rain4 <- group_by(rain, NAME)
rain5 <- summarize(rain4, meanrf = mean(PRCP, na.rm = TRUE))
# summarize makes a new column where its the mean of PRCP
head(rain5)
```

pipe operator (takes whats on the left as input for the right)
```{r}
rain5 <- rain %>% group_by(NAME) %>% summarize(meanrf = mean(PRCP, na.rm = TRUE))
head(rain5)
```

```{r}
####select
###create a tibble with only the Temp (e.g. those beginning with T) variables
###create another tibble that excludes latitude and longitude

mytibble <- select(rain, TAVG:TMIN)
#or
mytibble <- select(rain, TAVG, TMAX, TMIN)
#there are other ways, too. SEe the  transform data cheat sheet and look for helper functions for select. For example:
mytibble <- select(rain, starts_with("T"))
####mutate
######is elevation in feet or meters? Whichever it is, create a new variable in the new units.


```



Thursday

dates
- there are three classes for date variables
  - date class
  - POSIXct
  - POSIXit
- date classes store a date as the number since 1/1/1970
- CT stores seconds since 1/1/1970
- IT stores as a list, with separate elements for daat, mont, year, etc

```{r}
x <- as.Date('2022-03-31') # as date is always in THIS FORMAT
as.numeric(x) 
```

format
```{r}
?strptime #will give you all the format codes that you need to convert dates
d1 <- as.Date('03/31/22', format = '%m/%d/%Y') # your format has to be in quotes
# this is wrong becuase Y means uppcercase year and y is 2 digit year
d1 #incorrect
```

```{r}
d2 <- as.Date('03/31/22', format = '%m/%d/%y')
d2 #correct
```

practice
```{r}
d3 <- as.Date('03-31-2022', format = '%m-%d-%Y')
d4 <- as.Date('03/31/22', format = '%m/%d/%y')
d5 <- as.Date('March 31, 2022', format = '%B %d,%Y')
d3
d4
d5
```

try these
```{r}
library('lubridate')
d3 <- as.Date('03-31-2022', format = '%m-%d-%Y')
day(d3)
month(d3)
year(d3)
week(d3)
mday(d3)
wday(d3)
```


consider only LA county,
what dates are included?
how many neighborhoods?
do all neighborhoods have same number of obs
```{r}
covid <- read.csv('covid.csv')
head(covid)
# unique(covid$date)
# unique(covid$name)
covid2 <- covid %>% arrange(desc(date))
end <- covid2[1,'date']
covid3 <- covid %>% arrange(date)
start <- covid3[1,'date']
range <- c(start,end)
la <- covid %>% filter(county == "Los Angeles")
unique(covid$fips)
```

when you get a dataset the first steo is to generate questions about the data rather than 
blindly looking for patters

ggplot

- primary function is ggplot()
   - takes 2 arguments
    - data frame
    - aes
      - aesthetic that maps the variables in the data frame )position,size,color)
    - built in later
    
components 9arguments)
- date
- aesthetic mappings
- geo objects
- stat transforming
- scales
- faceting (subgroups)

p = ggplot(cdc #date, aes(weight,wtdesire) #columns) <- this gives an error bc u didnt give enough arguments like what type of geometric shape you want


p = ggplot(cdc #date, aes(weight,wtdesire) #columns) 
p + geom_point() #we want a scatter plot (THIS WORKS)


Doing more

p = ggplot(cdc #date, aes(weight,wtdesire) #columns) 
p + geom_point(aes(colour = gender)) # in scatterplot we are specifying using aes that we want the color of our points to be seperated by the gender of the point, it will give a different colour for each unique gender value



p = ggplot(cdc #date, aes(weight,wtdesire) #columns) 
p + geom_point(aes(colour = gender), size - 1, alpha = 0.75) #size changes the point size and alpha changes opacity, any changes that can be made to the point is in geom_point(aes
p+stat_smooth() #adds a line defaults to best fit curve
#many examples in the fuckinininin notES!!

```{r}
data("iris")
iris
?grep
?regexpr
version
vector <- c("c","meow")
```











