---
title: "Homework 2"
author: "Joshua Susanto"
date: "4/6/2022"
output: pdf_document
---

Please provide the following information in a pdf:

your answers and the process you used to get the answers (this should include the R code needed to do the calculations, and might also include words of explanation, assumptions made, etc.)

You can do this either through R markdown, or a word processing document (converted to pdf) that includes answers and explanations.

In either case, your answers should be easy to read and not require the reader to "hunt" for the correct values.

```{r}
library('tidyverse')
```


## 1) I took a random sample of people from the 2000 U.S. Census and recorded the incomes. The sample size was n = 30 (with replacement). The average of the sample was 23606 dollars and the standard deviation of the sample was $24757. Answer these questions using the statistics that I found from my sample.

### a) Find a 95% confidence interval for the mean income of the population.
```{r}
t <- qt(0.025, 29)
upper <- 23606 + (abs(t) * (24757/sqrt(30)))
lower <- 23606 - (abs(t) * (24757/sqrt(30)))
interval <- c(lower,upper)
print(interval)
```

### b) What assumptions must you make?
I assumed that the average of the sample was the mean of the sample. Due to the lack of information of population parameters, I also assumed that we are not using the CLT and thus are not under the assumption that the sampling distribution is approximately a normal distribution. 

### c) Your confidence interval is of the form (LB, UB) where LB and UB are both numbers. Suppose I'm going to take a new sample of size 30. Can I conclude that the probability the average of my sample is between LB and UB is 95%? (where LB and UB are the values you found in your interval.) Why or why not?

No we cannot assume this.
Our confidence interval does NOT reflect the probability that a different sample will fall into this interval. Rather it is a statistic that reflects that we are 95% CONFIDENT that the population parameter falls within our interval. Theoretically, as this interval was generated by only 1 sample, it is not reliable to say that another random sample will fall within this interval.

### d) A politician claims that the mean income of US residents (in 2000) was $25,000. Test this claim with an appropriate hypothesis test using a significance level of 5%.

Since we do not have the population standard deviation we will have to perform this hypothesis test using a t-test. The formula goes as follows:

$t = \frac{\overline{x} - \mu_o}{\frac{s}{\sqrt{n}}}$
```{r}
t <- (23606 - 25000)/(24757/sqrt(30))
2*pt(t,29)
```
Since the p value associated with our t score is 0.7599, which is greater than the significance level of 5%, we fail to reject the null hypothesis.

### e) For part (d), what's the smallest significance level you could have used so that the result would be that you would reject the null hypothesis?

The smallest significance level we could use is 76%. In this case, our p value would be slighty less than 0.76 and therefore we would reject the null hypothesis.


## 2) Suppose we've taken a random sample from the population consisting of all U.S. adults and administered COVID tests.Let p represent the proportion of the entire population that would have tested positive had we given them the same COVID test.  We find a 95% confidence interval for p.

### a) What happens to the width of the interval if we increase the sample size?

Since sample size is in the denominator of the confidence interval formula, if we increase the sample size the interval gets smaller.

### b) What happens to the width of the interval if we decrease the confidence level?

Since the confidence level is not in the deonominator of the confidence interval formula, if we decrease the confidence level the width of the interval will get smaller.


## 3) Referring to the cdc.csv data set (see the Module in Canvas that contains the data used in class), carry out a hypothesis test to test whether those who exercise (exerany==1) have a different mean desired weight than those who do not exercise. (hint: try the t.test function)

### a) state your null and alternative hypotheses in terms of symbols

Since we have 2 variables, we must do a hypothesis test as:

$H_0: \mu_1 - \mu_2 = 0$
$H_a: \mu_1 - \mu_2 \neq 0$

Where $\mu_1$ is the mean of those who exercised and $\mu_2$ is the mean of those who have not.

### b) give the value of your test statistic

```{r}
cdc <- read.csv('cdc.csv')
head(cdc)
exercise <- cdc %>% filter(exerany == 1)
noexercise <- cdc %>% filter(exerany == 0)
t.test(noexercise$wtdesire, exercise$wtdesire)
```
We get our test statistic to be -5.5844

### c) give the p-value of the test statistic
We get out p-value to be  2.415e-08


### d) state the decision you make, using a significance level of 5%

Using a significance value of 5%, we know that since our p-value is lower than  the significance value, we reject the null hypothesis and conclude that there is a significant difference between these 2 means.

### e) You read in the paper that someone says a p-value measures whether or not the null hypothesis is true.  State whether this is correct or incorrect. If correct, explain why.  If incorrect, give the correct meaning.   

This is incorrect. We are already under the assumption that the null hypothesis is true before performing our test. Rather, our hypothesis test and p-value will give us an indication as to whether or not to reject the null hypothesis (no longer assume it to be true) or keep the null (keep your initial assumption. The significance level is the probability that we are using to decide whether our statistic is too extreme.

### f) explain the meaning of the significance level to a lay person. That is, explain the definition without using technical jargon.

When we want to check if a result we get from an experiment or observation is likely due to random chance or something greater we use a hypothesis test to check this. We have our null hypothesis (the value we should theoretically be getting) and our alternate hypothesis (the value we got). We will measure the likelihood of getting our alternate hypothesis under the assumption that we should have got the null hypothesis, and our significance level is the cut-off. If the probability we got (called the p-value) is greater than our significance level then we can assume that there is no reason to suspect anything greater than random chance and we keep our initial assumption. However if the likelihood is lower than what our cut-off level is (meaning a very low chance) we reject our initial assumption and conclude that there is likely a greater reason for our test result.

