---
title: "1o1aweek3"
author: "Joshua Susanto"
date: "4/12/2022"
output: html_document
---

# two sample t test

H0: myu1 - myu2 = 0
Ha: myu 1 - mmyu2 =/= 0

to estimate difference between 2 means:
    - average 1 - average 2
    - X1 - X2
    
if null is true
    - E(X1 - X2) = 0
    
if observations are independent
      var(X1 - X2) = var(X1) + var(X2)
      
      var(X1) = sigma1^2 / n1
      
      var(X2) = sigma2^2 / n2
      
test stat might be:
    
      (X1 - X2)/√(var(X1) + var(X2))
      
hte null tells us about the means but not their values
it also doesnt tell us the values of SD nor their relationship

but if the SDs are equal

sigma1 == sigma2 == sigma

then you can pool all your data tgt and get a good estimate of sigma

s^2 pooled = (n1 - 1)s1^2 + (n2 - 1)s2^2
/n1 * n2 - 2

s^2 = (1(n-1)) * sum from i = 1 to n (xi - myu)^2

# pooled t stat

t = X1 - X2/√s^2pooled (i/n1 + 1/n2)

and this follows a t distribution with n1 + n2-2 degrees of freedom in R

t.test(x,y, alternative = 'two sided')


- if ur right and the sigmas are equal, then this is a good test
- if ur wrong then its a bad test. the SDs were incorrectly estimates and ur p val is wrong
- so in general it is safer to not assume equal SDs
- if they are not equal, you have a better estimator of the p value. if they are equal then ur right

if variances are unequal, we must estimate then seperately and use

      t = idk
      
this follows a t distributionwith dof approximated by

      df = idk
      
      t.test(x,y, alternative = 'two sided')
      
      
# all the prior stuff assumes that group 1 was independent of group 2
- what if you weight ppl at the start of a diet and again at the end and test the change in mean
- in that case this is NOT INDEPENDENT
- this is called the PAIRED T TEST
- t.test(x,y, alternative = ..., paired = TRUE)


        





```{r}
time <- c(0,3,60,100,140,180,210,265,330,390)
depth <- c(12.5,11.25,10.5,9.25,8,7,6,4.75,3.5,2.5)
plot(depth ~ time)
```
- looks linear
- decreasing 
```{r}
lemon <- lm(depth ~ time)
lemon_res <- resid(lemon)
plot(lemon_res)
abline(0,0)
```



```{r}
armspan <- read.csv('Armspan 2022 raw distribute.csv')
head(armspan)
str(armspan)
plot(armspan$What.is.your.height.in.INCHES..Note..5.feet...60.inches..6.feet...72.inches. ~ armspan$What.is.your.armspan.in.INCHES...Armspan.is.measured.as.the.greatest.distance.from.the.tip.of.the.fingers.of.the.left.hand.to.the.tip.of.the.fingers.of.the.right.when.your.arms.are.extended..parallel.to.the.ground.)
```








      